Supplementary Information

### 2021-08-26 doc/manuscript/ForestGEO_treerings.Rmd
fixes from USGS editor

\newpage
**Table 2 | Sites included in this analysis**. Here and throughout, sites are ordered by descending mean annual temperature. Additional site information is provided in Appendix `r site_details_appendix` and Table `r site_details`, and tree species and sampling details are detailed in Tables `r species_table`-`r sampling_details_table`. 
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(flextable)
set_flextable_defaults(font.family = "Arial", font.size = 8)
sites_table<- read.csv("tables_figures/sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
sites_table <- subset(sites_table, select = -c(MAT, JulyTsd, JanTsd, MAPsd)) #drop MAT column, plus columns with sd
ft <- flextable(sites_table)
ft <- autofit(ft)
ft <- width(ft, 1, width = .5)
ft <- width(ft, 2:3, width = 1)
ft <- width(ft, 4:6, width = .45)
ft <- width(ft, 7, width = .66)
ft <- width(ft, 8, width = .52)
ft <- width(ft, 9, width = .5)
ft <- width(ft, 10, width = 1.5)
ft <- footnote( ft, i = 1, j = 4:6,
                value = as_paragraph(c("Refers to 1950-2019 mean climate")),
                ref_symbols = c("*"),
                part = "header",inline=T)
ft <- footnote( ft, i = 1, j = 7,
                value = as_paragraph(c("BD= Broadleaf Deciduous, BE= Broadleaf Evergreen, C=Conifer")),
                ref_symbols = c("**"),
                part = "header",inline=T)
ft <- footnote( ft, i = c(6,8,9), j = 2,
                value = as_paragraph(c("Older forest, with majority of sampled trees established before 1900")),
                ref_symbols = c("#"),
                part = "body",inline=T)
ft

@@

We used the *climwin* package in R [@vandepol_identifying_2016] to identify the most influential climate variable (*i.e.*, that most strongly correlated with annual growth) and the seasonal window over which its effect was strongest for each of two categories of variables: a temperature group ($T_{min}$, $T_{max}$, $T_{mean}$, and PET) and a precipitation group (PPT, PDF).
To remove low-frequency variation that most likely represents responses to non-climatic drivers (*e.g.*, growth and aging of the tree, change in competitive dynamics, atmospheric pollution), we detrended the response variables by fitting penalized thin-plate regression splines in generalized additive models [GAM, functions *gam* and *s* in the R package *mgcv*, @wood_fast_2011] to individual growth records (RW, BAI, or $\Delta$AGB) from each core, and extracted the residual variation for each observation. 
The smoothing parameters were automatically selected by the *gam* function with generalized cross-validation (GCV).
Our application of the thin-plate regression splines acts similar to more traditional a priori detrending methods using a two-thirds spline commonly used in dendrochronology studies and results in similar predictor variable selection [Appendix `r traditional_comparison`, @cook_calculating_1997; @rollinson_climate_2021].  
We then used *climwin* to identify the climate drivers that most strongly correlated with the individual tree-level residuals of the growth variables, RW, BAI, or $\Delta$AGB, specifying linear and quadratic terms to allow for potential nonlinearities in the climate response.
Within *climwin*, we specified a mixed-effects model in which the fixed effects were the climate variables and the random intercepts were species (when n $\ge$ 3) and core identity (noting that these effects should be minimal given that residuals are centered around zero).
For each climate variable, we ran permutations for all possible combinations of consecutive months within a 15-month period ending near the time of cessation of formation of each annual ring (Table `r site_details`). 
*Climwin* runs all potential models to select the best fit (lowest AIC), and does k-fold cross-validation in its computation of AIC to guard against over-fitting [@vandepol_identifying_2016]. 
For each group of candidate climate variables (water and temperature; Fig. `r schematic`), we selected the variable - seasonal window combination with the lowest AIC as a candidate climate variable for the multivariate models.

>>>>

\newpage
**Table 2 | Sites included in this analysis**. Here and throughout, sites are ordered by descending mean annual temperature. Additional site information is provided in Appendix `r site_details_appendix` and Table `r site_details`, and tree species and sampling details are detailed in Tables `r species_table`-`r sampling_details_table`. 
```{r eval = TRUE, echo=FALSE, warning=FALSE}
library(flextable)
set_flextable_defaults(font.family = "Arial", font.size = 8)
sites_table<- read.csv("tables_figures/sites.csv", stringsAsFactors = FALSE, check.names = FALSE)
sites_table <- subset(sites_table, select = -c(MAT, JulyTsd, JanTsd, MAPsd)) #drop MAT column, plus columns with sd
ft <- flextable(sites_table)
ft <- autofit(ft)
ft <- width(ft, 1, width = .5)
ft <- width(ft, 2:3, width = 1)
ft <- width(ft, 4:6, width = .45)
ft <- width(ft, 7, width = .66)
ft <- width(ft, 8, width = .52)
ft <- width(ft, 9, width = .5)
ft <- width(ft, 10, width = 1.5)
ft <- footnote( ft, i = 1, j = 4:6,
                value = as_paragraph(c("Refers to 1950-2019 mean climate; T= mean monthly temperature; MAP= mean annual precipitation.")),
                ref_symbols = c("*"),
                part = "header",inline=T)
ft <- footnote( ft, i = 1, j = 7,
                value = as_paragraph(c("BD= Broadleaf Deciduous, BE= Broadleaf Evergreen, C=Conifer")),
                ref_symbols = c("**"),
                part = "header",inline=T)
ft <- footnote( ft, i = c(6,8,9), j = 2,
                value = as_paragraph(c("Older forest, with majority of sampled trees established before 1900")),
                ref_symbols = c("#"),
                part = "body",inline=T)
ft

@@

We used the *climwin* package in R [@vandepol_identifying_2016] to identify the most influential climate variable (*i.e.*, that most strongly correlated with annual growth) and the seasonal window over which its effect was strongest for each of two categories of variables: a temperature group ($T_{min}$, $T_{max}$, $T_{mean}$, and PET) and a precipitation group (PPT, PDF).
To remove low-frequency variation that most likely represents responses to non-climatic drivers (*e.g.*, growth and aging of the tree, change in competitive dynamics, atmospheric pollution), we detrended the response variables by fitting penalized thin-plate regression splines in generalized additive models [GAM, functions *gam* and *s* in the R package *mgcv*, @wood_fast_2011] to individual growth records (RW, BAI, or $\Delta$AGB) from each core, and extracted the residual variation for each observation. 
The smoothing parameters were automatically selected by the *gam* function with generalized cross-validation (GCV).
Our application of the thin-plate regression splines acts similar to more traditional a priori detrending methods using a two-thirds spline commonly used in dendrochronology studies and results in similar predictor variable selection [Appendix `r traditional_comparison`, @cook_calculating_1997; @rollinson_climate_2021].  
We then used *climwin* to identify the climate drivers that most strongly correlated with the individual tree-level residuals of the growth variables, RW, BAI, or $\Delta$AGB, specifying linear and quadratic terms to allow for potential nonlinearities in the climate response.
Within *climwin*, we specified a mixed-effects model in which the fixed effects were the climate variables and the random intercepts were species (when n $\ge$ 3) and core identity (noting that these effects should be minimal given that residuals are centered around zero).
For each climate variable, we ran permutations for all possible combinations of consecutive months within a 15-month period ending near the time of cessation of formation of each annual ring (Table `r site_details`). 
*Climwin* runs all potential models to select the best fit (lowest Akaike information criterion corrected for small sample size, AICc), and does k-fold cross-validation in its computation of AIC to guard against over-fitting [@vandepol_identifying_2016]. 
For each group of candidate climate variables (water and temperature; Fig. `r schematic`), we selected the variable - seasonal window combination with the lowest AIC as a candidate climate variable for the multivariate models.
